---
title: "Analysis"
author: "Alex Kan -lexokan"
date: "January 5, 2018"
output: html_document
---
```{r}
# Load packages
library(tidyverse)
library(RColorBrewer)
```

```{r}


general <- read_csv("temp/General.csv")
municipal <- read_csv("temp/Municipal.csv")

# Create data.frames of unqiue voters in each type of election
genUnique <- general[!duplicated(general[,"fullName"]),]
muniUnique <- municipal[!duplicated(municipal[,"fullName"]),]

# Plot histogram of general voters by age and status code. Note issues with inactive young voters.
ggplot(genUnique, aes(x = age, fill = statusCode)) +
    geom_histogram(bins = 20, color = 'black') +
    labs(x = 'Age', y = 'Count', title = 'General election voters by registration type')

# Plot histogram of municipal voters and age by status code. Once again not the lack of active young voters. Also note the much older median age. 
ggplot(muniUnique, aes(age, fill = statusCode)) +
    geom_histogram(bins = 20, color = 'black')+
    scale_fill_manual(values = c("blue","red","green")) +
    labs(x = 'Age', y = 'Voter Count', title = "Municipal Election Voters by Age")
  
  
```

```{r}
# Determine the age  in each Town during each election
genUnique %>%
    filter(municipalityName == "CARRBORO") %>%
    select(age) %>%
    lapply(., mean, na.rm = T)

genUnique %>%
    filter(municipalityName == "CHAPEL HILL") %>%
    select(age) %>%
    lapply(., mean, na.rm = T)
# Carrboro: 43, Chapel Hill: 38.5

muniUnique %>%
    filter(municipalityName == "CARRBORO") %>%
    select(age) %>%
    lapply(., mean, na.rm = T)

muniUnique %>% 
    filter(municipalityName == "CHAPEL HILL") %>% 
    select(age) %>%
    lapply(., mean, na.rm = T)
# Carrboro: 52, Chapel HillL: 57


# Determine the of registered voters who voted over the age of age 65 compared to 18-34 
muniUnique %>%
    select(age) %>%
    filter(. >= 65) %>%
    nrow(.)

muniUnique %>% 
    select(age) %>% 
    filter(.>= 18 & .<= 34) %>%
    nrow(.)
```
Nearly 2.5x as many Over 65 voters voted in the municipal election than the general 

```{r}
genUnique %>%
    select(age) %>%
    filter(. >= 65) %>%
    nrow(.)

genUnique %>%
    select(age) %>%
    filter(.>= 18 & .<= 34) %>%
    nrow(.)
```
4.0x as many under 34 voters voted than over 65 in the general 

```{r}
# Average age of active voters compared to average of all registered voters
combined %>%
    filter(statusCode == "A") %>% 
    select(age) %>%
    lapply(., mean, na.rm = T)

combined %>%
    filter(statusCode == "I") %>% 
    select(age) %>% 
    lapply(., mean, na.rm = T)

# Determine the number of voters of each type. 
filter(combined, statusCode == "A") %>%
    nrow()
filter(combined, statusCode == "I") %>%
    nrow()
filter(combined, statusCode == "S") %>%
    nrow()

```


```{r}
# Define function to determine the parcitipation rate by race of an election

participationRate <- function(df, municipality) {

# Create vector of race codes 
races <- c("A", "B", "M", "I", "O","W")

# Group by race and count the number of observations 
output <- df %>% 
    filter(municipalityName == municipality, race %in% races) %>%
    group_by(race) %>%
    summarise(
        count = n()
    )

# Create vector of population of each race within Chapel Hill
popByRace <- c(5785, 4630, 185, 905, 2650, 35075)

# Format output 
output[1] <- c("Asian", "Black", "Native American", "Multi-Racial", "Other/Unknown", "White")
output[3] <- (output[2] / popByRace)
output <- rename(output, participationRate = `count.1`)

list <- list(race = output[1], participationRate =  output[3])
output

}
```

Note the massive disparity in participation rates for muncipal elections, with an almost 30% difference between White voters and both Asian and African-American voters.
```{r}
participationRate(municipal, "CHAPEL HILL")

```


NOTES ON DATA -----------------------------------------------------------


 Population numbers from the 2011-2015 American Community Survey 5 year estimates - published 02/01/2017
 https://www.census.gov/rdo/data/voting_age_population_by_citizenship_and_race_cvap.html

 Voting age population in the Town of Chapel Hill by race

 Total		49230
 American Indian or Alaska Native Alone	170
 Asian Alone	5785
 Black or African American 4630
 Native Hawaiian or Other Pacific Islander Alone		15
 White Alone		35075
 American Indian or Alaska Native and White	85
 Asian and White	490
 Black or African American and White		165
 American Indian or Alaska Native and Black or African American	25
 Remainder of Two or More Race Responses		140



